DATA 621 - Business Analytics and Data Mining

### 2020-09-24 HW2/HW2-Group2.rmd
Finished section 10

10. ROC Curve
The receiver operating curve (ROC) is defined as a plot of the true positive
rate (TPR) against the false positive rate (FPR) at different probability
thresholds. The FPR is usually plotted on the x-axis. The function below takes
as input a data frame with columns named `class` and `pred_class` as above, and
a third column named `pred_prob` which reflects the predicted probability of
being positive. It then steps along a sequence from 0 to 1, assigning classes
based on that sequence, calculating the TPR and FPR for each threshold, plotting
the results when completed. Lastly, it calculates the area under the ROC (AUC).

```{r ROC}
ROC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    rocTable <- data.table(thresh = seq(0, 1, 0.01),
                           x = double(101),
                           y = double(101))
    for (i in seq_along(rocTable$thresh)) {
      dftmp <- data.frame(class = df$class,
                          pred_class = ifelse(df$pred_prob > rocTable$thresh[i],
                                              1, 0))
      rocTable$x[i] <- 1 - SPC(dftmp)
      rocTable$y[i] <- SNS(dftmp)
    }
    plot(rocTable$x, rocTable$y, type = 'l', xlab = "False Positive Rate",
         ylab = 'True Positive Rate')
  }
}
```

>>>>

10. ROC Curve
The receiver operating curve (ROC) is defined as a plot of the true positive
rate (TPR) against the false positive rate (FPR) at different probability
thresholds. The FPR is usually plotted on the x-axis. The function below takes
as input a data frame with columns named `class` and `pred_class` as above, and
a third column named `pred_prob` which reflects the predicted probability of
being positive. It then steps along a sequence from 0 to 1, assigning classes
based on that sequence, calculating the TPR and FPR for each threshold, plotting
the results when completed. Lastly, it calculates the area under the ROC (AUC).
Trapezoidal quadrature will be used to estimate the AUC from the table created
for the ROC, as the intervals on the x axis are not equidistant.

```{r ROC}
ROC <- function(df) {
  if (!is.data.frame(df)) {
    stop("Data set is not a dataframe. Try as.data.frame(dataset).")
  } else {
    rocTable <- data.frame(thresh = seq(0, 1, 0.01),
                           x = double(101),
                           y = double(101))
    for (i in seq_along(rocTable$thresh)) {
      dftmp <- data.frame(class = df$class,
                          pred_class = ifelse(df$pred_prob > rocTable$thresh[i],
                                              1, 0))
      rocTable$x[i] <- 1 - SPC(dftmp)
      rocTable$y[i] <- SNS(dftmp)
    }
  }
  return(list(plot = {plot(rocTable$x, rocTable$y, type = 'l',
                          xlab = "False Positive Rate",
                          ylab = 'True Positive Rate')
    lines(c(0, 1), c(0, 1), lty = 2)},
              AUC = sum((rocTable$y[1:100] + rocTable$y[2:101]) * 0.5 *
                          (rocTable$x[1:100] - rocTable$x[2:101]))))
}
```

The ROC and AUC for the dataset provided is shown below:

```{r ROCdata, echo=FALSE, fig.height=6}
ROC(class_data)
```
